### 2013-03-25 doc/dissertation/document/chapters/mc_theory/mc_theory.tex
updating symmetric leakage section in mc theory

\subsubsection{Random Walk Length}
\label{subsubsec:walk_length}
With the eigenvalue derivations verified, we can go about setting up
an experiment to measure the length of the random walks generated by
the adjoint Neumann-Ulam solver. To do this, we again use a $100
\times 100$ square grid with $h=0.1$ and the absorption cross varied
from 0 to 100 while the scattering cross section was fixed at
unity. Three weight cutoff values of \sn{1}{-2}, \sn{1}{-4}, and
\sn{1}{-8} were used with 10,000 histories generated by a point source
of strength 1 in the center of the domain. For each of the histories,
the number of transitions made was tallied to provide an effective
value of $k$ for each history. This value was then averaged over all
histories to get a measured value of $k$ for the particular
operator. Figure~\ref{fig:measured_length} presents these measurements
as well as the analytic result computed by Eq~(\ref{eq:analytic_k}) as
a function of the iteration matrix spectral radius,
$\rho(\ve{H})$. FigureOn the right,
Figure~\ref{fig:measured_length_error} gives the relative error
between the predicted and observed results. We note good qualitative
agreement between the measured and analytic results. However, we
observe a larger relative error for both long and short random walks.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/walk_length.png}
    \end{center}
    \caption{\textbf{Measured and analytic random walk length as a
        function of the iteration matrix spectral radius.} \textit{The
        weight cutoff was varied with \sn{1}{-2}, \sn{1}{-4}, and
        \sn{1}{-8}. The red data was computed numerically by an
        adjoint Neumann-Ulam implementation while the black dashed
        data was generated by Eq~(\ref{eq:analytic_k}).}}
    \label{fig:measured_length}
  \end{spacing}
\end{figure}
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/walk_length_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic random walk length relative
        error as a function of the iteration matrix spectral radius.}
      \textit{The weight cutoff was varied with \sn{1}{-2},
        \sn{1}{-4}, and \sn{1}{-8}. The relative error between the
        predicted and measured results is presented for each weight
        cutoff.}}
    \label{fig:measured_length_error}
  \end{spacing}
\end{figure}

@@

To compute the leakage fraction numerically, \sn{3}{5} histories were
sampled from a uniform source of strength unity over the global
domain. At the start of a stage of histories, the number of histories
starting in the center domain was computed and as the stage
progressed, the number of histories that exited that domain was
tallied with the ratio of the two numbers providing a numerical
measure for the leakage fraction. Figure~\ref{fig:measured_leakage}
gives the domain leakage measurements for the domain in the center of
the global grid as well as the analytic result computed by
Eqs~(\ref{eq:wigner_domain_leakage}) and
(\ref{eq:mean_chord_domain_leakage}) as a function of the iteration
matrix spectral radius.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=4.5in,clip]{chapters/mc_theory/leakage_variation.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage as a
        function of the iteration matrix spectral radius.} \textit{To
        test the behavior with respect to domain size, $n_i=50$,
        $n_i=100$,and $n_i=200$ were used. The red data was computed
        numerically by a domain-decomposed adjoint Neumann-Ulam
        implementation, the black dashed data was generated by
        Eq~(\ref{eq:mean_chord_domain_leakage}) using the mean-chord
        approximation, and the dashed-dotted black data was generated
        by Eq~(\ref{eq:wigner_domain_leakage}) using the Wigner
        rational approximation.}}
    \label{fig:measured_leakage}
  \end{spacing}
\end{figure}
Again, we note good qualitative agreement between the measured and
analytic quantities but we begin to see the limits of the leakage
approximations. To compare the quality of the two approximations, the
absolute error between the computed leakage fraction and that
generated by the Wigner rational and mean chord approximations is
plotted in Figure~\ref{fig:leakage_error} for all domain sizes
tested. 
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=4.5in,clip]{chapters/mc_theory/leakage_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage absolute
        error as a function of the iteration matrix spectral radius.}
      \textit{To test the behavior with respect to domain size,
        $n_i=50$ (green), $n_i=100$ (blue), and $n_i=200$ (red) were
        used. The dashed lines represent the error using the Wigner
        rational approximation while the solid lines represent the
        error using the mean-chord approximation.}}
    \label{fig:leakage_error}
  \end{spacing}
\end{figure}
From these error results, the mean chord approximation is shown to
have a lower error for ill-conditioned systems as compared to the
Wigner approximation while the Wigner approximation produces less
error for more well-conditioned systems. We also note that for the
optically thick domains, the error is likely corresponded to that
observed in Figure~\ref{fig:measured_length} for the $k$ parameter
while the large relative error in $k$ for optically thin domains does
not affect the approximation significantly. In general, the mean chord
approximation is a better choice to estimate the leakage fraction in a
domain from the adjoint Neumann-Ulam method and except for a single
data point with $n_i=50$, the mean chord approximation yielded leakage
fractions within 0.05 of the measured results. As the domain becomes
more optically thick (with both increasing $n_i$ and decreasing
$\rho(\ve{H})$), the approximations are more accurate.

>>>>

\subsubsection{Random Walk Length}
\label{subsubsec:walk_length}
With the eigenvalue derivations verified, we can go about setting up
an experiment to measure the length of the random walks generated by
the adjoint Neumann-Ulam solver. To do this, we again use a $100
\times 100$ square grid with $h=0.1$ and the absorption cross varied
from 0 to 100 while the scattering cross section was fixed at
unity. Three weight cutoff values of \sn{1}{-2}, \sn{1}{-4}, and
\sn{1}{-8} were used with 10,000 histories generated by a point source
of strength 1 in the center of the domain. For each of the histories,
the number of transitions made was tallied to provide an effective
value of $k$ for each history. This value was then averaged over all
histories to get a measured value of $k$ for the particular
operator. Figure~\ref{fig:measured_length} presents these measurements
as well as the analytic result computed by Eq~(\ref{eq:analytic_k}) as
a function of the iteration matrix spectral radius,
$\rho(\ve{H})$. Figure~\ref{fig:measured_length_error} gives the
relative error between the predicted and observed results. We note
good qualitative agreement between the measured and analytic
results. However, we observe a larger relative error for both long and
short random walks.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/walk_length.png}
    \end{center}
    \caption{\textbf{Measured and analytic random walk length as a
        function of the iteration matrix spectral radius.} \textit{The
        weight cutoff was varied with \sn{1}{-2}, \sn{1}{-4}, and
        \sn{1}{-8}. The red data was computed numerically by an
        adjoint Neumann-Ulam implementation while the black dashed
        data was generated by Eq~(\ref{eq:analytic_k}).}}
    \label{fig:measured_length}
  \end{spacing}
\end{figure}
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/walk_length_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic random walk length relative
        error as a function of the iteration matrix spectral radius.}
      \textit{The weight cutoff was varied with \sn{1}{-2},
        \sn{1}{-4}, and \sn{1}{-8}. The relative error between the
        predicted and measured results is presented for each weight
        cutoff.}}
    \label{fig:measured_length_error}
  \end{spacing}
\end{figure}

@@

To compute the leakage fraction numerically, \sn{3}{5} histories were
sampled from a uniform source of strength unity over the global
domain. At the start of a stage of histories, the number of histories
starting in the center domain was computed and as the stage
progressed, the number of histories that exited that domain was
tallied with the ratio of the two numbers providing a numerical
measure for the leakage fraction. Figure~\ref{fig:measured_leakage}
gives the domain leakage measurements for the domain in the center of
the global grid as well as the analytic result computed by
Eqs~(\ref{eq:wigner_domain_leakage}) and
(\ref{eq:mean_chord_domain_leakage}) as a function of the iteration
matrix spectral radius.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5.25in,clip]{chapters/mc_theory/leakage_variation.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage as a
        function of the iteration matrix spectral radius.} \textit{To
        test the behavior with respect to domain size, $n_i=50$,
        $n_i=100$,and $n_i=200$ were used. The red data was computed
        numerically by a domain-decomposed adjoint Neumann-Ulam
        implementation, the black dashed data was generated by
        Eq~(\ref{eq:mean_chord_domain_leakage}) using the mean-chord
        approximation, and the dashed-dotted black data was generated
        by Eq~(\ref{eq:wigner_domain_leakage}) using the Wigner
        rational approximation.}}
    \label{fig:measured_leakage}
  \end{spacing}
\end{figure}
Again, we note good qualitative agreement between the measured and
analytic quantities but we begin to see the limits of the leakage
approximations. To compare the quality of the two approximations, the
absolute error between the computed leakage fraction and that
generated by the Wigner rational and mean chord approximations is
plotted in Figure~\ref{fig:leakage_error} for all domain sizes
tested. 
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/leakage_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage absolute
        error as a function of the iteration matrix spectral radius.}
      \textit{To test the behavior with respect to domain size,
        $n_i=50$ (green), $n_i=100$ (blue), and $n_i=200$ (red) were
        used. The dashed lines represent the error using the Wigner
        rational approximation while the solid lines represent the
        error using the mean-chord approximation.}}
    \label{fig:leakage_error}
  \end{spacing}
\end{figure}
From these error results, the mean chord approximation is shown to
have a lower error for ill-conditioned systems as compared to the
Wigner approximation while the Wigner approximation produces less
error for more well-conditioned systems. We also note that for the
optically thick domains, the error is likely corresponded to that
observed in Figure~\ref{fig:measured_length} for the $k$ parameter
while the large relative error in $k$ for optically thin domains does
not affect the approximation significantly. In general, the mean chord
approximation is a better choice to estimate the leakage fraction in a
domain from the adjoint Neumann-Ulam method and except for a single
data point with $n_i=50$, the mean chord approximation yielded leakage
fractions within 0.05 of the measured results. As the domain becomes
more optically thick (with both increasing $n_i$ and decreasing
$\rho(\ve{H})$), the approximations are more accurate.
